---
title: "esm_206_lab5_km"
author: "Keene Morrow"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to aprse dates with `lubridate`
- Get counts of observations with count()
- Uncount() rows
- One- and two-sample t-tests
- Create heatmap with geom_tile()

```{r, include = FALSE}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>%
  janitor::clean_names()

```
Use tidyr::uncount() to convert lobster data from frequency format to case format
(aka, taking the lobster_count column and replicating each row the correct number of times to have one row per lobster found
lobster_count = 0 results in no rows of that data.  If those were necessary, complete_cases could be used)

```{r}
lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)
```

#### Exploraty Data Visualization

Considering site as variable of interest

```{r}
ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site),
                     color = "black") +
  facet_wrap(~site, scales = "free")
# scales = "free" removes the identical scale across the facets (can be used to make something ireesponsible)

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() + # default comparison is to normal distribution, but there are plenty of other options if needed
  facet_wrap(~site)
```

#### Convert 'date' column to class 'Data'

We'll use the `lubridate` package to convert to date format and then to parse month and year
```{r}

lobster_date <- lobster_tidy %>%
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Then we'll use the lubridate::month and lubridate::year functions to to parse the month and year

```{r}
lobster_parse_date <- lobster_date %>%
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

```

Now Let's find counts of observed lobstars based on different hierarchical groupings:

First, let's count lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <- lobster_parse_date %>%
  dplyr::count(obs_year)

lobster_y

lobster_site <- lobster_parse_date %>%
  dplyr::count(site)

lobster_site

```

If we want to create a sumamry table that contains statistics other than count by group, it's easier to use group_by() + n()


```{r}
lobster_summary <- lobster_parse_date %>%
  group_by(site) %>%
  summarize(
    lobster_number = n(),
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE)
  )

lobster_summary
```

#### Find confidence intervals

Use t.test() function to find confidenc eitnerval (for one sample) and perform t-tests to compate means of two samples

```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>%
  pull(size_mm) # pulls out vector of one variable

t.test(ivee_lobsters)

```
#### 2 sample t-test to compare means:
Is there a significant difference between lobster sizes observed at Naples reef versus lobsters observed at Mohawk reef? If they're from the same population and perfect smaples, the difference will be zero.  Follow up: are the means different enough to support whether or not the samples are from two different populations?

We've done our necessary exploratory analyses (look to be normally distributed, large n) in order to determine that a 2-sample t-test for means comaprison is appropriate.

```{r}
napl_sample <- lobster_tidy %>%
  filter(site == "NAPL") %>%
  pull(size_mm)

mohk_sample <- lobster_tidy %>%
  filter(site == "MOHK") %>%
  pull(size_mm)

mn_ttest <- t.test(napl_sample, mohk_sample)
mn_ttest
```

There is a significant difference in lobster lengths between Naples and Mohawk Reefs. (t(df) = statistic, p < 0.001, alpha = 0.05).

```{r}
lobster_mn <- lobster_tidy %>%
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn) # ~ = as a function of... different levels in the site column

mn_ttest2
```

There is a significant difference in lobster lengths between Naples and Mohawk Reefs. (t(df) = `r round(mn_ttest2$statistic,2)`, p < 0.001, alpha = 0.05).

#### Heatmap!

```{r}

lobster_ys <- lobster_parse_date %>%
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

